Preparación de datos

### 2025-01-10 04-analisis.qmd
add income per capita

# generate analytical sample
df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo, covid_economia, covid_beneficios, inc_pcmed) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
         )

@@

# Hipotesis
h0 <- "ola"
h1 <- "ola_num+ola_2"
h2 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo"
h3 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id"
h4 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia"
h5 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios"
h6 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc"
h7 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent"
h8 <- "ola_num+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"

@@

omit <- "(edad)|(ess)|(sexo)|(pos_id)"
texreg::knitreg(list(educ0, educ1, educ3, educ4, educ5, educ6,educ7,educ8),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05.
Note: Model 3 to 8 are controlled by age, gender, political position, and subjective social status.", 
                omit.coef = omit,
                custom.coef.names = c("Intercept",
                                      "Wave 2017",
                                      "Wave 2018",
                                      "Wave 2019",
                                      "Wave 2022",
                                      "Wave 2023",
                                      "Wave",
                                      "Wave^2",
                                      "Universitary (Ref.= Less than universitary)",
                                      "Income",
                                      # "High School",
                                      # "Technical",
                                      # "Universitary",
                                      # "Quintile 2",
                                      # "Quintile 3",
                                      # "Quintile 4",
                                      # "Quintile 5",
                                      # "Quintile (No information)",
#                                      "Subjective Social Status",
                                      # "Age 30-49",
                                      # "Age 50-64",
                                      # "Age 65 or more",
                                      # "Gender (Ref. Male)",
                                      # "Center",
                                      # "Right",
                                      # "Does not identify",
                                      "Covid: economy more important than health",
                                      "Covid: receive state benefits",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort"),
                groups = list("Wave (Ref.= Wave 2016)" = 2:6
                          #    "Education (Ref.= Primary)" = 9:11,
                  #"Household Income (Ref.= Quintile 1)" = 12:16
                  #"Age (Ref.= 18-29)" = 14:16,
                  #"Pol. pos (Ref.= Left)" = 18:20
                  ),
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.gof.rows = list("Controls"= c("No", "No", rep("Yes", 6))),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = FALSE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)

```

@@

h6 <- "ola+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"
h1b <- "ola+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_effort*inc_pcmed"
h3b <- "ola+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_talent*inc_pcmed"
h4b <- "ola+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+des_perc*inc_pcmed"
h5b <- "ola+ola_2+education+inc_pcmed+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+inc_pcmed*ola"
# A. justification educational inequality
educ1 <- lmer(formula(paste0("just_educ~",h1b,"+(inc_pcmed+merit_effort|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3b,"+(inc_pcmed+merit_talent|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4b,"+(inc_pcmed+des_perc|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5b,"+(ola+inc_pcmed|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

# table 3: interacciones
omit <- "(education)|(ess)|(edad)|(sexo)|(pos_id)|(covid_beneficios)|(covid_economia)"
texreg::knitreg(list(educ6,educ3,educ1, educ4, educ5),
                custom.note = "\\item *** p < 0.001; ** p < 0.01; * p < 0.05. Note: All the models are controlled by educational level, income quintile, subjective social status, age, gender, political position, and inequality perception",
                omit.coef = omit,
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.model.names = c("Model 8", "Model 9", "Model 10", "Model 11", "Model 12"),
                custom.coef.names = c("Intercept",
                                      "Wave",
                                      "Wave^2",
                                      "Income",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Income * Merit: Talent",
                                      "Income * Merit: Effort",
                                      "Income * Inequality perception",
                                      "Income * Wave"
                                      ),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual",
                                     "Var: Individuals Income",
                                     "Var: Individuals Merit: Talent",
                                     "Cov: Individuals (Intercept) Income",
                                     "Cov: Individuals (Intercept) Merit: Talent",
                                     "Cov: Individuals Income Merit: Talent",
                                     "Var: Individuals Merit: Effort",
                                     "Cov: Individuals (Intercept) Merit: Effort",
                                     "Cov: Individuals Income Merit: Effort",
                                     "Var: Individuals Inequality perception",
                                     "Cov: Individuals (Intercept) Inequality perception",
                                     "Cov: Individuals Income Inequality perception",
                                     "Var: Individuals Wave",
                                     "Cov: Individuals (Intercept) Wave",
                                     "Cov: Individuals Wave Income"),  
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = TRUE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)
```

>>>>

# generate analytical sample
df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo, covid_economia, covid_beneficios, ing_pc) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo),
         ing_pc_rec = (ing_pc/100000)
         )

@@

# Hipotesis
h0 <- "ola"
h1 <- "ola_num+ola_2"
h2 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo"
h3 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id"
h4 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia"
h5 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios"
h6 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc"
h7 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent"
h8 <- "ola_num+ola_2+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"

@@

omit <- "(edad)|(ess)|(sexo)|(pos_id)"
texreg::knitreg(list(educ0, educ1, educ3, educ4, educ5, educ6,educ7,educ8),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05.
Note: Model 3 to 8 are controlled by age, gender, political position, and subjective social status.", 
                omit.coef = omit,
                custom.coef.names = c("Intercept",
                                      "Wave 2017",
                                      "Wave 2018",
                                      "Wave 2019",
                                      "Wave 2022",
                                      "Wave 2023",
                                      "Wave",
                                      "Wave^2",
                                      "Universitary (Ref.= Less than universitary)",
                                      "Income per capita",
                                      # "High School",
                                      # "Technical",
                                      # "Universitary",
                                      # "Quintile 2",
                                      # "Quintile 3",
                                      # "Quintile 4",
                                      # "Quintile 5",
                                      # "Quintile (No information)",
#                                      "Subjective Social Status",
                                      # "Age 30-49",
                                      # "Age 50-64",
                                      # "Age 65 or more",
                                      # "Gender (Ref. Male)",
                                      # "Center",
                                      # "Right",
                                      # "Does not identify",
                                      "Covid: economy more important than health",
                                      "Covid: receive state benefits",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort"),
                groups = list("Wave (Ref.= Wave 2016)" = 2:6
                          #    "Education (Ref.= Primary)" = 9:11,
                  #"Household Income (Ref.= Quintile 1)" = 12:16
                  #"Age (Ref.= 18-29)" = 14:16,
                  #"Pol. pos (Ref.= Left)" = 18:20
                  ),
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.gof.rows = list("Controls"= c("No", "No", rep("Yes", 6))),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = FALSE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)
```

@@

h6 <- "ola+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort"
h1b <- "ola+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_effort*ola"
h3b <- "ola+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+merit_talent*ola"
h4b <- "ola+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+des_perc*ola"
h5b <- "ola+education+ing_pc_rec+ess+edad+sexo+pos_id+covid_economia+covid_beneficios+des_perc+merit_talent+merit_effort+ing_pc_rec*ola"
# A. justification educational inequality
educ1 <- lmer(formula(paste0("just_educ~",h1b,"+(ola+merit_effort|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3b,"+(ola+merit_talent|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4b,"+(ola+des_perc|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5b,"+(ola+ing_pc_rec|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

# table 3: interacciones
omit <- "(education)|(ess)|(edad)|(sexo)|(pos_id)|(covid_beneficios)|(covid_economia)"
texreg::knitreg(list(educ6,educ3,educ1, educ4, educ5),
                custom.note = "\\item *** p < 0.001; ** p < 0.01; * p < 0.05. Note: All the models are controlled by educational level, income quintile, subjective social status, age, gender, political position, and inequality perception",
                omit.coef = omit,
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.model.names = c("Model 8", "Model 9", "Model 10", "Model 11", "Model 12"),
                custom.coef.names = c("Intercept",
                                      "Wave",
                                      "Income",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Wave * Merit: Talent",
                                      "Wave * Merit: Effort",
                                      "Wave * Inequality perception",
                                      "Wave * Income"
                                      ),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual",
                                     "Var: Individuals Wave",
                                     "Var: Individuals Merit: Talent",
                                     "Cov: Individuals (Intercept) Wave",
                                     "Cov: Individuals (Intercept) Merit: Talent",
                                     "Cov: Individuals Wave Merit: Talent",
                                     "Var: Individuals Merit: Effort",
                                     "Cov: Individuals (Intercept) Merit: Effort",
                                     "Cov: Individuals Wave Merit: Effort",
                                     "Var: Individuals Inequality perception",
                                     "Cov: Individuals (Intercept) Inequality perception",
                                     "Cov: Individuals Wave Inequality perception",
                                     "Var: Individuals Income",
                                     "Cov: Individuals (Intercept) Income",
                                     "Cov: Individuals Wave Income"),  
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = TRUE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)
```

### 2025-01-10 production/prod_prep-mergedata.Rmd
add income per capita

df_study1_long_comunas <- 
 dplyr::left_join(x = df_study1_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

@@

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp, 
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

@@

df_study_long_t7_comunas <- 
 dplyr::left_join(x = df_study1_long_t7,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

>>>>

df_study1_long_comunas <- 
 dplyr::left_join(x = df_study1_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp, cambio_ingreso_pandemia, ing_pc,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

@@

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp, cambio_ingreso_pandemia, ing_pc,
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

@@

df_study_long_t7_comunas <- 
 dplyr::left_join(x = df_study1_long_t7,
                  y = df_study1_comunas, 
                  by = "comunacod") %>% 
 dplyr::select(idencuesta:edad,sexo, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp,cambio_ingreso_pandemia, ing_pc, 
               esc_cit17) %>% 
 dplyr::rename(comuna=comuna.x)

### 2025-01-10 production/prod_prep.Rmd
add income per capita

```{r}
# Reshape long to wide
df_study1_wide <- df_study1_long %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1_long,comuna,comunacod,ponderador_long_total,segmento, estrato, just_educ:edad,sexo, quintil1, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp))
                     )

@@

@@

```{r}
# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_",
                                 "^percentile_wave_", "^bot_mid_top_wave_", "^inc_pcmed_", "^ratioincome_", "^incomegrp_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo",
                             "percentile_wave", "bot_mid_top_wave", "inc_pcmed", "ratioincome", "incomegrp"),
              na.rm = F,value.factor = TRUE
              )
```

>>>>

```{r}
## Cambio entre primera ola con información y última ola con información
df_study1_long <- df_study1_long %>%
  group_by(idencuesta) %>% 
  mutate(
    ingreso_primera_ola = ing_pc[which.min(ola)],  # Ingreso en la primera ola
    ingreso_ultima_ola = ing_pc[which.max(ola)]   # Ingreso en la última ola
  ) %>%
  mutate(
    cambio_ingreso = case_when(
      is.na(ingreso_primera_ola) | is.na(ingreso_ultima_ola) ~ "NA",  # Si alguno es NA, categoría NA
      ingreso_ultima_ola > ingreso_primera_ola ~ "Mayor",
      ingreso_ultima_ola < ingreso_primera_ola ~ "Menor",
      ingreso_ultima_ola == ingreso_primera_ola ~ "Igual"
    )
  ) %>%
  ungroup()

df_study1_long$cambio_ingreso <- factor(df_study1_long$cambio_ingreso, levels = c("Menor", "Igual", "Mayor", "NA"))
df_study1_long$cambio_ingreso <- as.numeric(df_study1_long$cambio_ingreso)
sjmisc::frq(df_study1_long$cambio_ingreso)
```

```{r}
### Cambio pre y post pandemia
df_study1_long <- df_study1_long %>%
  group_by(idencuesta) %>%
  mutate(
    promedio_prepandemia = mean(ing_pc[ola %in% 1:4], na.rm = TRUE),  # Promedio de las olas 1 a 4
    promedio_postpandemia = mean(ing_pc[ola %in% 6:7], na.rm = TRUE)       # Promedio de las olas 6 y 7
  ) %>%
  mutate(
    cambio_ingreso_pandemia = case_when(
      is.na(promedio_prepandemia) | is.na(promedio_postpandemia) ~ NA_character_,  # Si faltan datos, NA
      promedio_postpandemia > promedio_prepandemia ~ "Mayor",
      promedio_postpandemia < promedio_prepandemia ~ "Menor",
      promedio_postpandemia == promedio_prepandemia ~ "Igual"
    )
  ) %>%
  ungroup()

df_study1_long$cambio_ingreso_pandemia <- factor(df_study1_long$cambio_ingreso_pandemia, levels = c("Menor", "Igual", "Mayor"))
sjmisc::frq(df_study1_long$cambio_ingreso_pandemia)
```

```{r}
# Reshape long to wide
df_study1_wide <- df_study1_long %>% 
  tidyr::pivot_wider(id_cols = c("idencuesta","muestra"),
                     names_from = "ola",
                     # names_prefix = c("educ","educyear","ing_pc","quintil","quintil1"),
                     values_from = names(select(df_study1_long,comuna,comunacod,ponderador_long_total,segmento, estrato, just_educ:edad,sexo, quintil1, ing_pc, percentile_wave, bot_mid_top_wave, inc_pcmed, ratioincome, incomegrp, cambio_ingreso, cambio_ingreso_pandemia))
                     )

@@

df_study1_wide$ing_pc_2 <-df_study1_wide$ing_pc_1 #income per capita
df_study1_wide$ing_pc_3 <-df_study1_wide$ing_pc_1
df_study1_wide$ing_pc_4 <-df_study1_wide$ing_pc_1
df_study1_wide$ing_pc_5 <-df_study1_wide$ing_pc_1
df_study1_wide$ing_pc_6 <-df_study1_wide$ing_pc_1
df_study1_wide$ing_pc_7 <-df_study1_wide$ing_pc_1

## Alternative incomes
df_study1_wide$percentile_wave_2 <- df_study1_wide$percentile_wave_1
df_study1_wide$percentile_wave_3 <- df_study1_wide$percentile_wave_1
df_study1_wide$percentile_wave_4 <- df_study1_wide$percentile_wave_1
df_study1_wide$percentile_wave_5 <- df_study1_wide$percentile_wave_1
df_study1_wide$percentile_wave_6 <- df_study1_wide$percentile_wave_1
df_study1_wide$percentile_wave_7 <- df_study1_wide$percentile_wave_1

df_study1_wide$bot_mid_top_wave_2 <- df_study1_wide$bot_mid_top_wave_1
df_study1_wide$bot_mid_top_wave_3 <- df_study1_wide$bot_mid_top_wave_1
df_study1_wide$bot_mid_top_wave_4 <- df_study1_wide$bot_mid_top_wave_1
df_study1_wide$bot_mid_top_wave_5 <- df_study1_wide$bot_mid_top_wave_1
df_study1_wide$bot_mid_top_wave_6 <- df_study1_wide$bot_mid_top_wave_1
df_study1_wide$bot_mid_top_wave_7 <- df_study1_wide$bot_mid_top_wave_1

df_study1_wide$inc_pcmed_2 <- df_study1_wide$inc_pcmed_1
df_study1_wide$inc_pcmed_3 <- df_study1_wide$inc_pcmed_1
df_study1_wide$inc_pcmed_4 <- df_study1_wide$inc_pcmed_1
df_study1_wide$inc_pcmed_5 <- df_study1_wide$inc_pcmed_1
df_study1_wide$inc_pcmed_6 <- df_study1_wide$inc_pcmed_1
df_study1_wide$inc_pcmed_7 <- df_study1_wide$inc_pcmed_1

df_study1_wide$ratioincome_2 <- df_study1_wide$ratioincome_1
df_study1_wide$ratioincome_3 <- df_study1_wide$ratioincome_1
df_study1_wide$ratioincome_4 <- df_study1_wide$ratioincome_1
df_study1_wide$ratioincome_5 <- df_study1_wide$ratioincome_1
df_study1_wide$ratioincome_6 <- df_study1_wide$ratioincome_1
df_study1_wide$ratioincome_7 <- df_study1_wide$ratioincome_1

df_study1_wide$incomegrp_2 <- df_study1_wide$incomegrp_1
df_study1_wide$incomegrp_3 <- df_study1_wide$incomegrp_1
df_study1_wide$incomegrp_4 <- df_study1_wide$incomegrp_1
df_study1_wide$incomegrp_5 <- df_study1_wide$incomegrp_1
df_study1_wide$incomegrp_6 <- df_study1_wide$incomegrp_1
df_study1_wide$incomegrp_7 <- df_study1_wide$incomegrp_1

@@

```{r}
# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_", "^edad_","^sexo_",
                                 "^percentile_wave_", "^bot_mid_top_wave_", "^inc_pcmed_", "^ratioincome_", "^incomegrp_",
                                 "^ing_pc_", "^cambio_ingreso_pandemia_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","quintil1","ess",
                             "pos_id", "edad","sexo",
                             "percentile_wave", "bot_mid_top_wave", "inc_pcmed", "ratioincome", "incomegrp",
                             "ing_pc", "cambio_ingreso_pandemia"),
              na.rm = F,value.factor = TRUE
              )
```
